### 2014-09-03 paper/paper.Rmd
albania is weird

@@

@@

```{r truancyschool, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
#school2012$name = school2012$name
school2012.sub = school2012 %>% dplyr::select(name, SCHOOLID, SC22Q01)
student2012.sub = student2012 %>% dplyr::select(name, SCHOOLID, math, read, sci, OECD)
#truancy.df = melt( student2012.sub %>% 
#  group_by(name, SCHOOLID) %>% 
#  left_join(school2012.sub, by = c("name", "SCHOOLID")) %>% 
#  group_by(name, SC22Q01, OECD) %>% summarise(math = mean(math), read = mean(read), sci = mean(sci)), 
#  id = c("name", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
truancy.df = melt( student2012.sub %>% 
  group_by(name, SCHOOLID) %>% 
  left_join(school2012.sub, by = c("name", "SCHOOLID")) %>% 
  group_by(name, SC22Q01, OECD) %>% summarise(math = mean(math)), 
  id = c("name", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
#truancy.df$group = paste0(truancy.df$name, truancy.df$variable)

>>>>

Albania appears to be an outlier in many educational trends.  Many socioeconomic and school factors that play a role in other nations'test scores do not appear to affect Albania's test scores.  For example, Albania has one of the smallest gender gaps for math and reading scores.  Additionally number of possessions and truancy rates affected test scores for every nation except Albania.

@@

Generally students with more possessions are expected to perform better in math, reading and science.  The number of possessions is a stand in for a student's economic standing.  The survey records whether students have desks, computers, internet, books, a place to study and dozens of other possessions.  After aggregating total number of possessions together, the figure ?*? shows that increasing number of possessions is correlated with higher test scores in all nations except Albania.

@@

A similar aberration is seen with school reported truancy rates.  Each school reports the level of student truancy: not at all, very little, to some extent, or a lot.  We group schools together by reported truancy level and aggregate the test scores for students at these schools.  Generally students at schools with high truancy rates average lower test scores than students at schools with lower truancy rates except for schools in Albania.

```{r truancyschool, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
#school2012$name = school2012$name
school2012.sub = school2012 %>% dplyr::select(name, SCHOOLID, SC22Q01)
student2012.sub = student2012 %>% dplyr::select(name, SCHOOLID, math, read, sci, OECD)
#truancy.df = melt( student2012.sub %>% 
#  group_by(name, SCHOOLID) %>% 
#  left_join(school2012.sub, by = c("name", "SCHOOLID")) %>% 
#  group_by(name, SC22Q01, OECD) %>% summarise(math = mean(math), read = mean(read), sci = mean(sci)), 
#  id = c("name", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
truancy.df = melt( student2012.sub %>% 
  group_by(name, SCHOOLID) %>% 
  left_join(school2012.sub, by = c("name", "SCHOOLID")) %>% 
  group_by(name, SC22Q01, OECD) %>% summarise(math = mean(math)), 
  id = c("name", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
levels(truancy.df$SC22Q01) = c("Not", "Little", "Some", "Lot")

#truancy.df$group = paste0(truancy.df$name, truancy.df$variable)
